---
title: "QMB Exercise 1 - Exploring Housing Rents"
author: "Dino Nienhold"
date: "19 Feb, 2015"
output: pdf_document
---

#Introduction
The following report is based on the QMB Exercise 1 - Exploring Housing Rents. The task description pdf file is bis_ex1-HousingRents.pdf

#Requirements
Please make sure that you the following packages loaded in your workspace.
```{r}
library("dplyr")
library("ggplot2")
```

#Data Set
Please make sure you have the file housingrents.csv in the subdirectoy Data in your workspace.
```{r}
housingrents <- read.csv("./Data/housingrents.csv",sep=";")
```
\pagebreak

#Task 1
There are `r dim(housingrents)[1]` observations and `r dim(housingrents)[2]` variables in the dataset (Use `dim(housingrents)`).
The str command gives an overview of the variable types:
```{r}
str(housingrents)
```
There are 14 NA values in the balcony variable.
```{r}
summary(housingrents)
```
\pagebreak

#Task 2

##Data Processing
For analysis purposes it is necessary to convert the rooms and nre variable to a factor.
```{r}
housingrents <- mutate(housingrents, rooms = factor(rooms), 
  nre = factor(nre,levels=c(0,1),labels=c("no","yes")))
```

##Plotting
The following chart shows the frequency of appartments according to their numberof rooms.
```{r}
ggplot(data=housingrents, aes(x=rooms,label=rooms))  +
  geom_bar(fill="blue") + 
  ggtitle("Count by Rooms") 
```

\pagebreak

#Task 3
In this section the contingency table for rooms and nre are calculated and plotted.
```{r}
rooms2nre <- xtabs(~rooms+nre, data=housingrents)
rooms2nre <- prop.table(rooms2nre,1)
ggplot(data.frame(rooms2nre), aes(x=rooms, y=Freq, fill=nre)) + 
  geom_bar(stat="identity") + 
  xlab("Number of Rooms") + 
  ylab("Ratio of NRE Property to non-Property") + 
  ggtitle("NRE Property by Number of Rooms") 
```

Contingency Table with row percentages
```{r}
addmargins(prop.table(rooms2nre,1))
```

\pagebreak

#Task 4
In this section the distribution of the variable area is analyzed to see if it is normally distributed.
```{r}
#Calculate binwidth based on the Freedman-Diaconis rule
bw <- diff(range(housingrents$area)) / (2 * IQR(housingrents$area) /
  length(housingrents$area)^(1/3))
#Plot the histogram
g <- ggplot(housingrents, aes(x = area)) + 
  geom_histogram(binwidth = bw,colour="black", fill="white")  + 
  geom_density(alpha=.2) + 
  geom_density(alpha=.5, fill="#FF6666") +
  geom_vline(aes(xintercept=mean(area, na.rm=T)),
  color="red", linetype="dashed", size=1)
print(g)
```

\pagebreak

The qqplot should to get a better understanding.
```{r}
qqnorm(housingrents$area);qqline(housingrents$area, col = 2)
```

##Conclusion
The area variable is normally distributed.

\pagebreak

#Task 5
In this section the mean/median of the rent of NRE and non-NRE appartments are compared
```{r}

```

```{r}
ggplot(housingrents,aes(y=rent,x=nre, fill=nre)) +  
  geom_boxplot()  +
  stat_summary(fun.y=mean, geom="point", shape=1, size=4) + 
  xlab("Property owned by NRE") + 
  ylab("Amount of Rent") + 
  ggtitle("Rent Comparison of nre and non-nre appartments") 
```

The following table shows the mean and median rent by nre and non-nre appartments:
```{r}
select(housingrents,rent,nre) %>% 
  group_by(nre) %>% 
  summarise(mean = mean(rent), median = median(rent) )
```

\pagebreak

#Task 6
This sections shows that the variables rent and the area are associated.
```{r}
ggplot(housingrents, aes(x=area, y=rent)) +
    geom_point(shape=1) +  geom_smooth(method=lm) +
    ylim(0, 5000) + 
    xlim(0,300)

```
    
The outliers are appartmentswith a high rent >2500
```{r}
filter(housingrents, rent > 2500)
```
Without the outliers the correlation between the rent and the area stays around the same.
```{r}
ggplot(filter(housingrents, rent <= 2500 ), aes(x=area, y=rent)) +
     geom_point(shape=1) +  geom_smooth(method=lm) +
    ylim(0, 5000) + 
    xlim(0,300)
```

\pagebreak

#Task 7
In this section the rent per square meter by nre and non-nre apartments are analyzed.
```{r}
#Create new variable rps (rent per squre meter
housingrents <- mutate(housingrents,rps = rent/area)
ggplot(housingrents,aes(y=rps,x=nre, fill=nre)) +  
  geom_boxplot()  +
  stat_summary(fun.y=mean, geom="point", shape=1, size=4) + 
  xlab("Property owned by NRE") + 
  ylab("Rent per square meter") + 
  ggtitle("Rent per square meter Comparison of nre and non-nre appartments") 
```

The following table shows the mean and median rent per square meter by nre and non-nre appartments:
```{r}
select(housingrents,rps,nre) %>% 
  group_by(nre) %>% 
  summarise(mean = mean(rps), median = median(rps) )
```
##Conclusion
The table shows that nre are only slightly more expensive when measured by rent per square meter.
